---
title: "Final Independent Project"
author: "Marisa Mackie"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    toc: TRUE
    theme: architect
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction
This project is related to my Master's Thesis project:

**Mapping chemical stimuli to their cognate neurons in _Pristionchus pacificus_ via genetically encoded calcium indicators**


I work with nematode (microscopic roundworm) species _Pristionchus pacificus_ and am interested in studying their amphid neurons, which are responsible for the sensing of chemicals in their environment--that is, odorants and tastants. There are 12 pairs of these amphid neurons, however, it is unknown which neurons specifically correspond to particular volatile odors or water-soluble tastants. My goal is to map various chemical stimuli to their cognate neurons.

To do this, I have used genetically-encoded calcium indicators (GECIs) (i.e., GCaMP) to measure neuron response to chemical stimuli in real time. 
Neuron response is measured and ultimately calculated as **percent change in fluorescence intensity (df/f) over time**.

### Purpose:

I want to visualize df/f over time in the form of line plots, where any peaks along the line would indicate a large change in fluorescence (and therefore, neural activity). I would hope to see peaks that correspond to the times I administered the chemical stimulus to the worm.

I want to visualize:

* df/f for each individual worm

* Average df/f of each treatment group


## Libraries
These are the necessary libraries we will need to load.
```{r}
library(here)
library(tidyverse)
library(tidytext)
```

## Data
Data is stored in the **data** folder.
This data was collected and stored in a _.log_ file, in the form of 9 comma-delimited columns.
The data is organized in folders in this order:

_worm strain & stimulant_ --> _tracking Journal_ --> _worm ID_ --> .log file

During analysis, we will be using _for loops_ to iterate through each of these folders & grab the worm data.


## Functions
For the purposes of this project, I have created a handful of functions that we will need to use for data analysis:

**dff()**
The entire project is based around the measurement of percent change in fluorescence in the neuron over time. This is called _delta f / f_ (a.k.a _dff_).

Therefore, we require a function to calculate dff from our raw fluorescence data.

```{r}
## Function for calculating df/f (% change in fluorescence)
dff <- function(data) {
  # sets baseline fluorescence
  f0 <- mean(data[10:40]) # averages fluorescence data from frame 10-40 (1-4s)
  # calculates df/f as a percentage
  dff <- ((data - f0)/f0)*100
  # returns dff
  return(dff)
}
```


**e_to_num()**
This function is due to an issue I ran into while analyzing fluorescence data.
The fluorescence data is stored in scientific notation, e.g. _1.2345e+008_.
However, R reads this as a _string_ due to the "e" and "+", but we need R to read it as a _numeric_ so we can use it for mathematical calculations (i.e., calculating dff).

Therefore, we require a function to convert this character into a double. In doing so, we would convert it from scientific notation to its expanded form.

```{r}
## Function to convert strings of scientific notation containing "e" (i.e., 1.234e+005) to numeric in expanded form (double)
e_to_num <- function(string) {
  # finds first instance of "e" in the string, subsets everything before it & converts it to a numeric (double)
  value <- as.numeric(substring(string, 1, unlist(gregexpr("e", string))[1]))
  # finds first instance of "+" in the string, subsets everything after it & converts it to a numeric (double)
  # note: this number is the exponent, so we will set 10^exponent
  exponent <- 10^(as.numeric(str_sub(string, unlist(gregexpr("\\+", string))[1]+1, str_length(string))))
  # multiplies value by the 10^n exponent
  expanded_num <- value*exponent
  # returns numeric (double) of original string in expanded form
  return(expanded_num)
}
```


## Part 1
Part 1 has multiple objectives that will be split into sub-parts:

* **Iterate through the folders & get the data (.log files)**

* **Clean data & convert to clean .csv**

* **Calculate dff values for each individual worm**

* **Plot dff values for each individual worm**


### Part 1a - Iterate through folders & get the data
As previously mentioned, the data is stored in nested folders as such:

_worm strain & stimulant --> tracking Journal --> worm ID --> .log file_

Therefore, we need a for-loop to iterate through each of these.

The basic outline would look something like this:

```{r, warning = FALSE, message = FALSE}
## Iterates through each stimulant folder
#for (i in 1:length(stim_folders)){
  
  ## Iterates through each JNL folder
  #for (j in 1:length(jnl_folders)){
    
    ## Iterates through each worm folder
    #for (k in 1:length(worm_folders)){
      
    ## } # ends iteration through worm folders
    
  ## } # ends iteration through JNL folders
  
## } # ends iteration through stim_folders

```

So we set up the necessary paths for each one:

```{r, warning = FALSE, message = FALSE}
# points to worm strain + stimulant folder on the computer
stim_path <- here("IndependentProject_Mackie", "data", "GCaMP_data")
# lists all folders in that path
stim_folders <- dir(path = stim_path)

# Iterates through each stimulant folder
for (i in 1:length(stim_folders)){
  
  # points to JNL folder on the computer
  jnl_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i])
  # lists all JNL folders in that path
  jnl_folders <- dir(path = jnl_path)
  
  # Iterates through each JNL folder within the current stimulant folder
  for (j in 1:length(jnl_folders)){
    # points to worm folder on the computer
    worm_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i], jnl_folders[j])
    # lists all worm folders in that path
    worm_folders <- dir(path = worm_path)
    
    # Iterates through each worm folder within the current JNL folder
    for (k in 1:length(worm_folders)){
      
    } # ends iteration through worm folders
    
  } # ends iteration through JNL folders
  
} # ends iteration through stim_folders
```

Now that we have made it to the logfiles, we can read them in:

```{r, warning = FALSE, message = FALSE}
### within worm-folder iteration ###

# points to the location of tracked worm data (logfiles) on the computer
      logfile_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i], jnl_folders[j], worm_folders[k])
      # lists all logfiles in that path (contains ".log")
      logfiles <- dir(path = logfile_path, pattern = ".log")
      
      # reads in current logfile into rawdata
      rawdata <- read_log(here("IndependentProject_Mackie","data", "GCaMP_data", stim_folders[i],jnl_folders[j],worm_folders[k],logfiles))
```

Here's one example of the raw data from one of the .log files:

```{r}
glimpse(rawdata)
```

Now that the logfiles are being read in with each iteration, while we're here, we can begin to clean the data.

### Part 1b - Clean data & convert to clean .csv

It makes sense to do this while still in the for-loops because it will do the cleaning to **_each_** logfile.

Of the 9 initial columns, I only need to keep the first 3. I also decided to add 2 columns to give information about which treatment groups (Stimulant & JNL) that each worm belongs to.

```{r, warning = FALSE, message = FALSE}
### within worm-folder iteration ###

# reads in current logfile into rawdata
      rawdata <- read_log(here("IndependentProject_Mackie","data", "GCaMP_data", stim_folders[i],jnl_folders[j],worm_folders[k],logfiles))
      
      wormdata <- rawdata %>% 
        # adds column for stimulant and JNL type
        mutate(stim_jnl = paste(stim_folders[i],jnl_folders[j], sep = "_"),
               # renames default column names (Xn) to specified variable names
               frame = rawdata$X1,
               time_s = (rawdata$X2)/1000, # converts time in ms to s
               fluor = rawdata$X3,
               .keep = "used") # only keeps mutated columns
```

Here is an example of what one of our clean data looks like:

```{r}
glimpse(wormdata)
```


### Part 1c - Calculate dff values for each individual worm

Next, here is where I did my **calculations for dff** which required me to use the **dff()** function I showed earlier. 

However, I also ran into an issue of needing to convert strings to numerics before I could successfully calculate dff, which is why I also created the **e_to_num()** function.

```{r, warning = FALSE, message = FALSE}
### within worm-folder iteration ###

# For fluor column, removes commas and converts values from sci notation (chr) to expanded form (dbl)
      wormdata <- wormdata %>% 
        mutate(fluor = str_replace_all(fluor, "\\,",""),
               fluor = e_to_num(fluor)) %>% 
        # adds column for calculating dff
        mutate(dff = dff(fluor))
```

Here is where I then converted my cleaned & dff-calcuated data into a _.csv_ file and saved it to my **data folder** to use later.

Since this is still being iterated, it created one .csv file for _each_ logfile.

```{r, warning = FALSE, message = FALSE}
### within worm-folder iteration ###

# converts wormdata to csv and saves to clean_csv folder within data folder, naming it based on stimulant, JNL, and worm ID
      write_csv(wormdata, here("IndependentProject_Mackie", "data", "clean_csv", paste(stim_folders[i],jnl_folders[j],logfiles, sep = "_")))
```

### Part 1d - Plot dff values for each individual worm

Still within the iteration, I created plots of dff over time for _each individual worm_. 

```{r, warning = FALSE, message = FALSE}
### within worm-folder iteration ###

# generates plot of % df/f over time in s
      ggplot()+
        
        # axis scales by 10
        scale_x_continuous(breaks = seq(from = 0, to = 180, by=10))+
        
        # adds red vertical line to notable timepoints for stimulant ON & OFF
        geom_vline(xintercept = 10, color = "red")+
        geom_vline(xintercept = 130, color = "red")+
        
        # specifies line plot, & size/color of line
        geom_line(data = wormdata,
                  mapping = aes(x = time_s, y = dff),
                  color = "black", linewidth = 1)+
        
        # specifies plot labels
        labs(x = "Time (s)",
             y = "% dF/F",
             title = "Change in fluorescence over time",
             subtitle = paste(str_replace_all(stim_folders[i], "_", " "), jnl_folders[j]),
             caption = worm_folders[k])+
        
        # specifies color and size of plot elements
        theme(plot.title = element_text(size = 12, face = "bold"),
              plot.subtitle = element_text(size = 10, color = "saddlebrown"),
              plot.caption = element_text(size = 8, color = "gray30"),
              axis.title = element_text(size = 10))
      
      # saves plot to output folder, and names it by worm folder name
      ggsave(here("IndependentProject_Mackie","output","individual_dff",paste("plot",stim_folders[i],jnl_folders[j],worm_folders[k],".png", sep = "-")),
             width = 6, height = 6)
```




Lastly, I made sure to close each loop.


Here is _Part1_ all together:

```{r}
# pre-allocates a dataframe to hold information for convenience later
# 4 columns, but we will append more rows later
size_mat <- data.frame(matrix(ncol = 5))

# worm_num is the current worm being analyzed overall
worm_num = 0

# points to worm strain + stimulant folder on the computer
stim_path <- here("IndependentProject_Mackie", "data", "GCaMP_data")
# lists all folders in that path
stim_folders <- dir(path = stim_path)

# Iterates through each stimulant folder
for (i in 1:length(stim_folders)){
  
  # points to JNL folder on the computer
  jnl_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i])
  # lists all JNL folders in that path
  jnl_folders <- dir(path = jnl_path)
  
  # Iterates through each JNL folder within the current stimulant folder
  for (j in 1:length(jnl_folders)){
    # points to worm folder on the computer
    worm_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i], jnl_folders[j])
    # lists all worm folders in that path
    worm_folders <- dir(path = worm_path)
    
    # Iterates through each worm folder within the current JNL folder
    for (k in 1:length(worm_folders)){
      
      # adds +1 to worm count
      worm_num = worm_num + 1
      
      # points to the location of tracked worm data (logfiles) on the computer
      logfile_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i], jnl_folders[j], worm_folders[k])
      # lists all logfiles in that path (contains ".log")
      logfiles <- dir(path = logfile_path, pattern = ".log")
      
      # reads in current logfile into rawdata
      rawdata <- read_log(here("IndependentProject_Mackie","data", "GCaMP_data", stim_folders[i],jnl_folders[j],worm_folders[k],logfiles))
      
      wormdata <- rawdata %>% 
        # adds column for stimulant and JNL type
        mutate(stim_jnl = paste(stim_folders[i],jnl_folders[j], sep = "_"),
               # renames default column names (Xn) to specified variable names
               frame = rawdata$X1,
               time_s = (rawdata$X2)/1000, # converts time in ms to s
               fluor = rawdata$X3,
               .keep = "used") # only keeps mutated columns
      
      # For fluor column, removes commas and converts values from sci notation (chr) to expanded form (dbl)
      wormdata <- wormdata %>% 
        mutate(fluor = str_replace_all(fluor, "\\,",""),
               fluor = e_to_num(fluor)) %>% 
        # adds column for calculating dff
        mutate(dff = dff(fluor))
      
      # converts wormdata to csv and saves to clean_csv folder within data folder, naming it based on stimulant, JNL, and worm ID
      write_csv(wormdata, here("IndependentProject_Mackie", "data", "clean_csv", paste(stim_folders[i],jnl_folders[j],logfiles, sep = "_")))
      
      # generates plot of % df/f over time in s
      ggplot()+
        
        # axis scales by 10
        scale_x_continuous(breaks = seq(from = 0, to = 180, by=10))+
        
        # adds red vertical line to notable timepoints for stimulant ON & OFF
        geom_vline(xintercept = 10, color = "red")+
        geom_vline(xintercept = 130, color = "red")+
        
        # specifies line plot, & size/color of line
        geom_line(data = wormdata,
                  mapping = aes(x = time_s, y = dff),
                  color = "black", linewidth = 1)+
        
        # specifies plot labels
        labs(x = "Time (s)",
             y = "% dF/F",
             title = "Change in fluorescence over time",
             subtitle = paste(str_replace_all(stim_folders[i], "_", " "), jnl_folders[j]),
             caption = worm_folders[k])+
        
        
        # specifies color and size of plot elements
        theme(plot.title = element_text(size = 12, face = "bold"),
              plot.subtitle = element_text(size = 10, color = "saddlebrown"),
              plot.caption = element_text(size = 8, color = "gray30"),
              axis.title = element_text(size = 10))
      
      # saves plot to output folder, and names it by worm folder name
      ggsave(here("IndependentProject_Mackie","output","individual_dff",paste("plot",stim_folders[i],jnl_folders[j],worm_folders[k],".png", sep = "-")),
             width = 6, height = 6)
      
      # number of frames is the number of rows within wormdata
      frames <- nrow(wormdata)
      
      # dataframe to store time information (to use for plotting later)
      time_data <- wormdata$time_s
      
    } # ends iteration through worm folders
    
    # adds column names to size_mat
    colnames(size_mat) <- c("frames","stim_jnl","worms_jnl","worm_num","jnl_start")
    # appends rows to size_mat
    size_mat[nrow(size_mat) + 1,] = c(frames,
                                      paste(stim_folders[i],jnl_folders[j],sep = "_"), # stimulant + jnl
                                      k, # final worm of each journal (= worms per journal)
                                      worm_num, # worm count (= total worms up to that journal)
                                      worm_num - k + 1) # starting worm for each jnl
    
  } # ends iteration through JNL folders
  
} # ends iteration through stim_folders

# drop first row (Nas) from size_mat
size_mat <- drop_na(size_mat)

```

Although, during this iteration, I created some helpful variables to help me later on during analysis.

For example:

```{r}
glimpse(size_mat)
```

_____________________________________________________________________________

The next 3 parts will all be related to producing the **Average dff** plots.

For this, we will use the clean .csv files that we created!

This is the approach:

1. Gather all individual dff values from the csv files into a single dataframe called **dff_all**

2. **Average** dff values by Stimulant + JNL and store them in *avg_all*

3. **Plot Average dff!**

## Part 2 - Gather dff values into dff_all
We previously calculated dff for each worm and stored it into our clean csv files.

Therefore, to grab those dff values, we must iterate through each csv file.

```{r}

# points to location of csv file on the computer
csv_path <- here("IndependentProject_Mackie", "data", "clean_csv")
# lists all .csv files in that path
csvfiles <- dir(path = csv_path)

# pre-allocates a dataframe to store all dff values
dff_all <- data.frame(matrix(nrow = frames, ncol = length(csvfiles)))
# adds column names as dff{n}
colnames(dff_all) <- c(paste("dff_worm",(1:length(csvfiles)),sep = ""))

# Iterates through each csv file (i.e., for each worm)
for (l in 1:length(csvfiles)){
  
  # reads in csv files
  csv_data <- read_csv(here("IndependentProject_Mackie", "data", "clean_csv", csvfiles[l]))
  
  # stores dff column into dff_all
  dff_all[l] <- csv_data$dff
}

```

Here are all the dff values for each worm:

```{r}
glimpse(dff_all)
```


## Part 3 - Average dff values by Stimulant + Journal

Now that we have dff values all stored in dff_all, we can now take the averages of those based on the Stimulant + JNL combination.

To do this, we can use _size_mat_ to iterate through the Stimulant + JNLs (treatments), which only requires one for-loop.

_(As opposed to previously, where we iterated through 3 sets of folders and therefore needed 3 nested for-loops. Creating the .csv files was useful for reducing the amount of separate iterations we must do to access our data.)_

```{r}
# pre-allocates a dataframe to store all average dff values (averaged by journal)
# 2 columns because there are 2 unique stim + jnl combinations
avg_all <- data.frame(matrix(nrow = frames, ncol = nrow(size_mat)))
# adds column names as stim_jnl{n}
colnames(avg_all) <- c(paste("stim_jnl",(1:nrow(size_mat)),sep = ""))

# iterates over each row in size_mat (i.e., m = the current stim_jnl)
for (m in 1:nrow(size_mat)){
  
  # if first stim_jnl, starts with 1st worm
  ifelse(m == 1, start_worm <-  1,
         # else, starts with 1st worm of the current journal
         start_worm <- size_mat$jnl_start[m])
  
  # ending worm = final worm of current stim_jnl
  end_worm <- size_mat$worm_num[m]
  
  # calculates averages across rows, only for worms that correspond to the current stim_jnl (removes NAs)
  # stores this in column of avg_all corresponding to current stim_jnl
  avg_all[m] <- rowMeans(dff_all[,start_worm:end_worm])
  
} # ends iteration for each jnl (row in size_mat)
    
```

Here are the Average dff values (for each treatment group = Stimulant + JNL).
```{r}
glimpse(avg_all)
```



## Part 4 - Plot Average dff
Here is where the Average dff (per treatment group) was plotted.
To do this, we have once again used the handy variable _size_mat_ to help us iterate through the Stimulant + JNLs, without having to iterate through 3 different folders!

Note: This likely could have been done in the previous iteration, but I chose to do it separately for ease of understanding. 

```{r}
# pre-allocates dataframe to hold only the x & y values for plotting
plot_data <- data.frame(matrix(nrow = frames, ncol = 2))
# column names
colnames(plot_data) <- c("x","y")

# iterates for each journal (# of rows in size_mat)
for (n in 1:nrow(size_mat)){
  
  # updates plot_data with appropriate data
  plot_data <- plot_data %>% 
    mutate(x = time_data, # assigns time data to x column
           y = avg_all[,n]) # assigns current column of avg_all to y column
  
  # generates plot of average % df/f over time in s
  ggplot(plot_data)+
    
    # axis scales by 10
    scale_x_discrete(breaks = seq(from = 0, to = 180, by = 10))+
    
    # adds red vertical line to notable timepoints for stimulant ON & OFF
    geom_vline(xintercept = 10, color = "red")+
    geom_vline(xintercept = 130, color = "red")+
    
    # specifies line plot, & size/color of line
    geom_line(mapping = aes(x = x,
                            y = y),
              color = "blue", linewidth = 2)+
    
    # specifies plot labels
    labs(x = "Time (s)",
         y = "% dF/F",
         title = "Average Change in fluorescence over time",
         subtitle = size_mat$stim_jnl[n])+ # current stimulant + jnl
    
    
    # specifies color and size of plot elements
    theme(plot.title = element_text(size = 12, face = "bold"),
          plot.subtitle = element_text(size = 10, color = "saddlebrown"),
          plot.caption = element_text(size = 8, color = "gray30"),
          axis.title = element_text(size = 10))
  
  # saves plot to output folder, and names it by worm folder name
  ggsave(here("IndependentProject_Mackie","output","average_dff",paste("plot",size_mat$stim_jnl[n],".png", sep = "-")),
         width = 6, height = 6)
  
} # ends iteration
```

Here is one of the two Average dff plots I have so far! 
(One for each treatment group)

```{r, echo = FALSE}
# pre-allocates dataframe to hold only the x & y values for plotting
plot_data <- data.frame(matrix(nrow = frames, ncol = 2))
# column names
colnames(plot_data) <- c("x","y")

# iterates for each journal (# of rows in size_mat)
for (n in 1:nrow(size_mat)){
  
  # updates plot_data with appropriate data
  plot_data <- plot_data %>% 
    mutate(x = time_data, # assigns time data to x column
           y = avg_all[,n]) # assigns current column of avg_all to y column
  
  # generates plot of average % df/f over time in s
  avg_dff_plot <- ggplot(plot_data)+
    
    # axis scales by 10
    scale_x_discrete(breaks = seq(from = 0, to = 180, by = 10))+
    
    # adds red vertical line to notable timepoints for stimulant ON & OFF
    geom_vline(xintercept = 10, color = "red")+
    geom_vline(xintercept = 130, color = "red")+
    
    # specifies line plot, & size/color of line
    geom_line(mapping = aes(x = x,
                            y = y),
              color = "blue", linewidth = 2)+
    
    # specifies plot labels
    labs(x = "Time (s)",
         y = "% dF/F",
         title = "Average Change in fluorescence over time",
         subtitle = size_mat$stim_jnl[n])+ # current stimulant + jnl
    
    
    # specifies color and size of plot elements
    theme(plot.title = element_text(size = 12, face = "bold"),
          plot.subtitle = element_text(size = 10, color = "saddlebrown"),
          plot.caption = element_text(size = 8, color = "gray30"),
          axis.title = element_text(size = 10))
  
  # saves plot to output folder, and names it by worm folder name
  ggsave(here("IndependentProject_Mackie","output","average_dff",paste("plot",size_mat$stim_jnl[n],".png", sep = "-")),
         width = 6, height = 6)
  
} # ends iteration

avg_dff_plot
```


## All together now!
Here is the _entire_ project, where all of these parts work together to produce my desired data visualizations!

```{r}
#### Load Libraries ####-----------------------------------------------
library(tidyverse)
library(here)
library(tidytext)

#### Functions ####----------------------------------------------------

## Function to convert strings of scientific notation containing "e" (i.e., 1.234e+005) to numeric in expanded form (double)
e_to_num <- function(string) {
  # finds first instance of "e" in the string, subsets everything before it & converts it to a numeric (double)
  value <- as.numeric(substring(string, 1, unlist(gregexpr("e", string))[1]))
  # finds first instance of "+" in the string, subsets everything after it & converts it to a numeric (double)
  # note: this number is the exponent, so we will set 10^exponent
  exponent <- 10^(as.numeric(str_sub(string, unlist(gregexpr("\\+", string))[1]+1, str_length(string))))
  # multiplies value by the 10^n exponent
  expanded_num <- value*exponent
  # returns numeric (double) of original string in expanded form
  return(expanded_num)
}

## Function for calculating df/f (% change in fluorescence)
dff <- function(data) {
  # sets baseline fluorescence
  f0 <- mean(data[10:40]) # averages fluorescence data from frame 10-40 (1-4s)
  # calculates df/f as a percentage
  dff <- ((data - f0)/f0)*100
  # returns dff
  return(dff)
}

#### Analysis ####-----------------------------------------------------
#### Part 1 - Get data, convert to clean csv, calculate dff, plot individual dff ####----------

# pre-allocates a dataframe to hold information for convenience later
# 4 columns, but we will append more rows later
size_mat <- data.frame(matrix(ncol = 5))

# worm_num is the current worm being analyzed overall
worm_num = 0

# points to worm strain + stimulant folder on the computer
stim_path <- here("IndependentProject_Mackie", "data", "GCaMP_data")
# lists all folders in that path
stim_folders <- dir(path = stim_path)

# Iterates through each stimulant folder
for (i in 1:length(stim_folders)){
  
  # points to JNL folder on the computer
  jnl_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i])
  # lists all JNL folders in that path
  jnl_folders <- dir(path = jnl_path)
  
  # Iterates through each JNL folder within the current stimulant folder
  for (j in 1:length(jnl_folders)){
    # points to worm folder on the computer
    worm_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i], jnl_folders[j])
    # lists all worm folders in that path
    worm_folders <- dir(path = worm_path)
    
    # Iterates through each worm folder within the current JNL folder
    for (k in 1:length(worm_folders)){
      
      # adds +1 to worm count
      worm_num = worm_num + 1
      
      # points to the location of tracked worm data (logfiles) on the computer
      logfile_path <- here("IndependentProject_Mackie", "data", "GCaMP_data", stim_folders[i], jnl_folders[j], worm_folders[k])
      # lists all logfiles in that path (contains ".log")
      logfiles <- dir(path = logfile_path, pattern = ".log")
      
      # reads in current logfile into rawdata
      rawdata <- read_log(here("IndependentProject_Mackie","data", "GCaMP_data", stim_folders[i],jnl_folders[j],worm_folders[k],logfiles))
      
      wormdata <- rawdata %>% 
        # adds column for stimulant and JNL type
        mutate(stim_jnl = paste(stim_folders[i],jnl_folders[j], sep = "_"),
               # renames default column names (Xn) to specified variable names
               frame = rawdata$X1,
               time_s = (rawdata$X2)/1000, # converts time in ms to s
               fluor = rawdata$X3,
               .keep = "used") # only keeps mutated columns
      
      # For fluor column, removes commas and converts values from sci notation (chr) to expanded form (dbl)
      wormdata <- wormdata %>% 
        mutate(fluor = str_replace_all(fluor, "\\,",""),
               fluor = e_to_num(fluor)) %>% 
        # adds column for calculating dff
        mutate(dff = dff(fluor))
      
      # converts wormdata to csv and saves to clean_csv folder within data folder, naming it based on stimulant, JNL, and worm ID
      write_csv(wormdata, here("IndependentProject_Mackie", "data", "clean_csv", paste(stim_folders[i],jnl_folders[j],logfiles, sep = "_")))
      
      # generates plot of % df/f over time in s
      ggplot()+
        
        # axis scales by 10
        scale_x_continuous(breaks = seq(from = 0, to = 180, by=10))+
        
        # adds red vertical line to notable timepoints for stimulant ON & OFF
        geom_vline(xintercept = 10, color = "red")+
        geom_vline(xintercept = 130, color = "red")+
        
        # specifies line plot, & size/color of line
        geom_line(data = wormdata,
                  mapping = aes(x = time_s, y = dff),
                  color = "black", linewidth = 1)+
        
        # specifies plot labels
        labs(x = "Time (s)",
             y = "% dF/F",
             title = "Change in fluorescence over time",
             subtitle = paste(str_replace_all(stim_folders[i], "_", " "), jnl_folders[j]),
             caption = worm_folders[k])+
        
        
        # specifies color and size of plot elements
        theme(plot.title = element_text(size = 12, face = "bold"),
              plot.subtitle = element_text(size = 10, color = "saddlebrown"),
              plot.caption = element_text(size = 8, color = "gray30"),
              axis.title = element_text(size = 10))
      
      # saves plot to output folder, and names it by worm folder name
      ggsave(here("IndependentProject_Mackie","output","individual_dff",paste("plot",stim_folders[i],jnl_folders[j],worm_folders[k],".png", sep = "-")),
             width = 6, height = 6)
      
      # number of frames is the number of rows within wormdata
      frames <- nrow(wormdata)
      
      # dataframe to store time information (to use for plotting later)
      time_data <- wormdata$time_s
      
    } # ends iteration through worm folders
    
    # adds column names to size_mat
    colnames(size_mat) <- c("frames","stim_jnl","worms_jnl","worm_num","jnl_start")
    # appends rows to size_mat
    size_mat[nrow(size_mat) + 1,] = c(frames,
                                      paste(stim_folders[i],jnl_folders[j],sep = "_"), # stimulant + jnl
                                      k, # final worm of each journal (= worms per journal)
                                      worm_num, # worm count (= total worms up to that journal)
                                      worm_num - k + 1) # starting worm for each jnl
    
  } # ends iteration through JNL folders
  
} # ends iteration through stim_folders

# drop first row (Nas) from size_mat
size_mat <- drop_na(size_mat)

"Part 1 complete!
Clean csv files generated (see data)
Individual dff plots generated (see outputs)"

#### Part 2 - Gather dff values into dff_all ####-------------

# points to location of csv file on the computer
csv_path <- here("IndependentProject_Mackie", "data", "clean_csv")
# lists all .csv files in that path
csvfiles <- dir(path = csv_path)

# pre-allocates a dataframe to store all dff values
dff_all <- data.frame(matrix(nrow = frames, ncol = length(csvfiles)))
# adds column names as dff{n}
colnames(dff_all) <- c(paste("dff_worm",(1:length(csvfiles)),sep = ""))

# Iterates through each csv file (i.e., for each worm)
for (l in 1:length(csvfiles)){
  
  # reads in csv files
  csv_data <- read_csv(here("IndependentProject_Mackie", "data", "clean_csv", csvfiles[l]))
  
  # stores dff column into dff_all
  dff_all[l] <- csv_data$dff
}

"Part 2 complete!
Gathered all dff values into dff_all."


#### Part 3 - Average dff values by Stimulant + Journal & Plot them! ####-------------

# pre-allocates a dataframe to store all average dff values (averaged by journal)
# 2 columns because there are 2 unique stim + jnl combinations
avg_all <- data.frame(matrix(nrow = frames, ncol = nrow(size_mat)))
# adds column names as stim_jnl{n}
colnames(avg_all) <- c(paste("stim_jnl",(1:nrow(size_mat)),sep = ""))

# iterates over each row in size_mat (i.e., m = the current stim_jnl)
for (m in 1:nrow(size_mat)){
  
  # if first stim_jnl, starts with 1st worm
  ifelse(m == 1, start_worm <-  1,
         # else, starts with 1st worm of the current journal
         start_worm <- size_mat$jnl_start[m])
  
  # ending worm = final worm of current stim_jnl
  end_worm <- size_mat$worm_num[m]
  
  # calculates averages across rows, only for worms that correspond to the current stim_jnl (removes NAs)
  # stores this in column of avg_all corresponding to current stim_jnl
  avg_all[m] <- rowMeans(dff_all[,start_worm:end_worm])
  
} # ends iteration for each jnl (row in size_mat)

"Part 3 complete!
Average dff calculated and stored in avg_all"
    

#### Part 4 - Plotting Average dff ####-------------

# pre-allocates dataframe to hold only the x & y values for plotting
plot_data <- data.frame(matrix(nrow = frames, ncol = 2))
# column names
colnames(plot_data) <- c("x","y")

# iterates for each journal (# of rows in size_mat)
for (n in 1:nrow(size_mat)){
  
  # updates plot_data with appropriate data
  plot_data <- plot_data %>% 
    mutate(x = time_data, # assigns time data to x column
           y = avg_all[,n]) # assigns current column of avg_all to y column
  
  # generates plot of average % df/f over time in s
  ggplot(plot_data)+
    
    # axis scales by 10
    scale_x_discrete(breaks = seq(from = 0, to = 180, by = 10))+
    
    # adds red vertical line to notable timepoints for stimulant ON & OFF
    geom_vline(xintercept = 10, color = "red")+
    geom_vline(xintercept = 130, color = "red")+
    
    # specifies line plot, & size/color of line
    geom_line(mapping = aes(x = x,
                            y = y),
              color = "blue", linewidth = 2)+
    
    # specifies plot labels
    labs(x = "Time (s)",
         y = "% dF/F",
         title = "Average Change in fluorescence over time",
         subtitle = size_mat$stim_jnl[n])+ # current stimulant + jnl
    
    
    # specifies color and size of plot elements
    theme(plot.title = element_text(size = 12, face = "bold"),
          plot.subtitle = element_text(size = 10, color = "saddlebrown"),
          plot.caption = element_text(size = 8, color = "gray30"),
          axis.title = element_text(size = 10))
  
  # saves plot to output folder, and names it by worm folder name
  ggsave(here("IndependentProject_Mackie","output","average_dff",paste("plot",size_mat$stim_jnl[n],".png", sep = "-")),
         width = 6, height = 6)
  
} # ends iteration

"Part 4 complete!
Average dff plots generated! (see outputs)"

"Done!"
```


## Future notes
Although this does what I need and works pretty well, a project can always be improved upon!

Here are some things I'd like to work on & add to this project in the future:

* Plot individual worm dff & average dff together. This makes it easier to see how close individual worms are to each other, and how the average compares to each worm in the sample.

* Making the code "future-proof." I want to be able to add new data to it without worrying about my code breaking. The data I am working with have different journal programs, and therefore different numbers of frames and times. I would need to make adjustments so that my data can be plotted successfully for all of my different experimental conditions.

* Additional tidying & efficiency in general.

* Other calculations I could make & other ways to visualize my data.



###### Thanks for reading!